---
title: "Comparing Stokes to Rich"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
resource_files:
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(here)
library(plotly)
library(conflicted)
library(janitor)
library(readxl)
conflicts_prefer(dplyr::filter)

#functions------------------------------
row_to_string <- function(row) {
  map2_chr(names(row), row, ~ paste0(.x, ": ", .y)) |>
    paste(collapse = ", ")
}

get_cagr <- function(tbbl){
  max_year <- max(tbbl$year)
  min_year <- min(tbbl$year)
  start <- tbbl$value[tbbl$year==min_year]
  end <-  tbbl$value[tbbl$year==max_year]
  (end/start)^(1/(max_year-min_year))-1 
}

rich_vs_stokes <- function(tbbl, var1, var2, cor){
  plt <- ggplot(tbbl,  aes({{  var1  }},
                           {{  var2  }}))+
    geom_abline(slope = 1, intercept = 0, colour="white", lwd=2)+
    geom_line(mapping=aes(group=interaction(bc_region, lmo_detailed_industry)), lwd=.2, alpha=.2)+
    geom_point(mapping=aes(colour=year,
                           text=paste0(
                             "Region: ",
                             bc_region,
                             "\n Industry: ",
                             lmo_detailed_industry,
                             "\n Year: ",
                             year,
                             "\n Rich: ",
                             scales::comma({{  var1  }}, accuracy = 1),
                             "\n Stokes: ",
                             scales::comma({{  var2  }}, accuracy = 1))), size=.5)+
    scale_colour_viridis_c()+
    scale_x_continuous(trans="log10", labels=scales::comma)+
    scale_y_continuous(trans="log10", labels=scales::comma)+
    labs(title=paste("correlation: ",round(cor,3)))

  plotly::ggplotly(plt, tooltip="text")
}
rich_vs_stokes2 <- function(tbbl, var1, var2, cor){
  plt <- ggplot(tbbl,  aes({{  var1  }},
                           {{  var2  }}))+
    geom_abline(slope = 1, intercept = 0, colour="white", lwd=2)+
    geom_line(mapping=aes(group=description), lwd=.2, alpha=.2)+
    geom_point(mapping=aes(colour=year,
                           text=paste0(
                             "NOC: ",
                             description,
                             "\n Year: ",
                             year,
                             "\n Rich: ",
                             scales::comma({{  var1  }}, accuracy = 1),
                             "\n Stokes: ",
                             scales::comma({{  var2  }}, accuracy = 1))), size=.5)+
    scale_colour_viridis_c()+
    scale_x_continuous(trans="log10", labels=scales::comma)+
    scale_y_continuous(trans="log10", labels=scales::comma)+
    labs(title=paste("correlation: ",round(cor,3)))

  plotly::ggplotly(plt, tooltip="text")
}

demand_plot <- function(tbbl, rich, stokes, title){
  plt <- tbbl|>
  ggplot(aes(x={{  rich  }}, 
                           y={{  stokes  }}, 
                           text=paste0("Region: ", bc_region,
                                       "\n Industry: ", lmo_detailed_industry,
                                       "\n Rich: ", scales::comma({{  rich  }}, accuracy = 1),
                                       "\n Stokes: ", scales::comma({{  stokes  }}, accuracy = 1))))+
  geom_abline(slope=1, intercept=0, linewidth=2, color="white") +
  geom_point(alpha=.25)+
  scale_x_continuous(trans="log10", labels=scales::comma)+
  scale_y_continuous(trans="log10", labels=scales::comma)+
  labs(title=title)

plotly::ggplotly(plt, tooltip="text")
}

demand_plot2 <- function(tbbl, rich, stokes, title){
  plt <- tbbl|>
  ggplot(aes(x={{  rich  }}, 
                           y={{  stokes  }}, 
                           text=paste0("NOC: ", noc_5,
                                       "\n Description: ", description,
                                       "\n Rich: ", scales::comma({{  rich  }}, accuracy = 1),
                                       "\n Stokes: ", scales::comma({{  stokes  }}, accuracy = 1))))+
  geom_abline(slope=1, intercept=0, linewidth=2, color="white") +
  geom_point(alpha=.25)+
  scale_x_continuous(trans="log10", labels=scales::comma)+
  scale_y_continuous(trans="log10", labels=scales::comma)+
  labs(title=title)

plotly::ggplotly(plt, tooltip="text")
}



#read in the data----------------------
noc_names <- read_csv(here("data","mapping","noc21descriptions.csv"))|>
  select(noc_5=lmo_noc, description=class_title)|>
  distinct()|>
  mutate(noc_5=str_pad(noc_5, "0", side = "left", width=5))

correct_industries <- read_excel(here("data","mapping", "industry_mapping_2025_with_stokes_agg.xlsx"))|>
  select(lmo_ind_code, lmo_detailed_industry)|>
  distinct()|>
  ungroup()

census_occupation <- noc_names|>
  inner_join(read_rds(here("out", "census_occupation.rds")))

census_industry_region <- read_rds(here("out", "census_industry_region.rds"))|>
  select(-lmo_detailed_industry)|>
  inner_join(correct_industries)

#my forecasts-----------------------------------
richs_forecast <- read_rds(here("out","richs_forecast.rds"))

richs_region_industry <- richs_forecast|>
  group_by(bc_region, lmo_ind_code, lmo_detailed_industry, year)|>
  summarise(richs_employment=sum(employment, na.rm = TRUE),
            richs_expansion=sum(expansion_demand, na.rm = TRUE),
            richs_replacement=sum(replacement_demand, na.rm = TRUE))
            
richs_occupation <- richs_forecast|>
  group_by(noc_5, year)|>
  summarise(richs_employment=sum(employment, na.rm = TRUE),
            richs_expansion=sum(expansion_demand, na.rm = TRUE),
            richs_replacement=sum(replacement_demand, na.rm = TRUE))

correct_regions <- richs_region_industry|>
  ungroup()|>
  select(bc_region)|>
  distinct()

#lfs data-----------------------

occ_lfs <- read_rds(here("out","historic_noc_5.rds"))|>
  select(noc_5, year, value=employment)|>
  inner_join(noc_names)


ind_lfs_file <- here("data","Employment for 64 LMO Industries 2000-2024.xlsx")

ind_lfs <- excel_sheets(ind_lfs_file) %>%
  set_names() %>%
  map(~ read_excel(ind_lfs_file, sheet = .x, skip = 3))|>
  enframe()|>
  unnest(value)|>
  rename(bc_region=name)|>
  pivot_longer(cols=starts_with("2"), names_to = "year")|>
  clean_names()|>
  fuzzyjoin::stringdist_join(correct_industries)|>
  rename(lmo_detailed_industry=lmo_detailed_industry.y)|>
  fuzzyjoin::stringdist_join(correct_regions)|>
  rename(bc_region=bc_region.y)

#stokes data-------------------
stokes_employment_industry <- read_excel(here("data","stokes_data", "employment_industry.xlsx"), skip = 3)|>
  pivot_longer(cols=starts_with("2"), names_to = "year", values_to = "stokes_employment")|>
  clean_names()|>
  ungroup()|>
  rename(lmo_detailed_industry=industry)|>
  fuzzyjoin::stringdist_join(correct_industries|>select(lmo_detailed_industry))|>
  rename(lmo_detailed_industry=lmo_detailed_industry.y)|>
  select(bc_region=geographic_area, lmo_detailed_industry, year, stokes_employment)|>
  filter(bc_region!="British Columbia")|>
  mutate(year=as.numeric(year),
         bc_region=if_else(bc_region=="Mainland South West","Lower Mainland-Southwest", bc_region),
         bc_region=if_else(bc_region=="Vancouver Island Coast","Vancouver Island and Coast", bc_region)
  )|>
  fuzzyjoin::stringdist_full_join(correct_regions)|>
  rename(bc_region=bc_region.y)|>
  select(-bc_region.x)

stokes_employment_occupation <- read_excel(here("data","stokes_data", "employment_occupation.xlsx"), skip = 3)|>
  pivot_longer(cols=starts_with("2"), names_to = "year", values_to = "stokes_employment")|>
  clean_names()|>
  filter(geographic_area!="British Columbia")|>
  group_by(noc, year)|>
  summarize(stokes_employment=sum(stokes_employment))|>
  mutate(year=as.numeric(year),
         noc_5=str_sub(noc, 2))|>
  inner_join(noc_names)

stokes_jo_occupation <- read_excel(here("data","stokes_data", "job_openings_occupation.xlsx"), skip = 3)|>
  filter(`Geographic Area`=="British Columbia",
         Variable %in% c("Expansion Demand", "Replacement Demand"),
         NOC!="#T")|>
  select(NOC, Variable, starts_with("2"))|>
  pivot_longer(cols=starts_with("2"), names_to = "year", values_to = "stokes")|>
  pivot_wider(names_from = Variable, values_from = stokes)|>
  rename(stokes_expansion=`Expansion Demand`,
         stokes_replacement=`Replacement Demand`)|>
  mutate(noc_5=str_remove_all(NOC, "#"),
         year=as.numeric(year), .before=everything())|>
  select(-NOC)

stokes_jo_region_industry <- read_excel(here("data","stokes_data", "job_openings_industry.xlsx"), skip = 3)|>
  filter(Variable %in% c("Expansion Demand", "Replacement Demand"))|>
  pivot_longer(cols=starts_with("2"), names_to = "year", values_to = "stokes")|>
  clean_names()|>
  ungroup()|>
  rename(lmo_detailed_industry=industry)|>
  fuzzyjoin::stringdist_join(correct_industries|>select(lmo_detailed_industry))|>
  rename(lmo_detailed_industry=lmo_detailed_industry.y)|>
  select(bc_region=geographic_area, lmo_detailed_industry, year, variable, stokes)|>
  filter(bc_region!="British Columbia")|>
  mutate(year=as.numeric(year),
         bc_region=if_else(bc_region=="Mainland South West","Lower Mainland-Southwest", bc_region),
         bc_region=if_else(bc_region=="Vancouver Island Coast","Vancouver Island and Coast", bc_region)
  )|>
  fuzzyjoin::stringdist_full_join(correct_regions)|>
  rename(bc_region=bc_region.y)|>
  select(-bc_region.x)|>
  pivot_wider(names_from = variable, values_from = stokes)|>
  rename(stokes_expansion=`Expansion Demand`,
         stokes_replacement=`Replacement Demand`)

#merging forecasts------------------------------

jo_region_industry <- full_join(richs_region_industry, stokes_jo_region_industry)|>
  filter(year>min(year))|>
  select(-richs_employment)|>
  group_by(bc_region, lmo_detailed_industry)|>
  summarize(richs_expansion=sum(richs_expansion, na.rm = TRUE),
            richs_replacement=sum(richs_replacement, na.rm = TRUE),
            stokes_expansion=sum(stokes_expansion, na.rm = TRUE),
            stokes_replacement=sum(stokes_replacement, na.rm = TRUE))
  
jo_occupation <- full_join(richs_occupation, stokes_jo_occupation)|>
  filter(year>min(year))|>
  select(-richs_employment)|>
  group_by(noc_5)|>
  summarize(richs_expansion=sum(richs_expansion, na.rm = TRUE),
            richs_replacement=sum(richs_replacement, na.rm = TRUE),
            stokes_expansion=sum(stokes_expansion, na.rm = TRUE),
            stokes_replacement=sum(stokes_replacement, na.rm = TRUE))|>
  left_join(noc_names)

employment_region_industry <- full_join(richs_region_industry, stokes_employment_industry)
employment_occupation <- full_join(richs_occupation, stokes_employment_occupation)

#correlations between forecasts----------------------------

industry_employment_cor <- cor(employment_region_industry$richs_employment, 
                               employment_region_industry$stokes_employment, 
                               method = "pearson", 
                               use="pairwise.complete.obs")

occupation_employment_cor <- cor(employment_occupation$richs_employment, 
                               employment_occupation$stokes_employment, 
                               method = "pearson",
                               use="pairwise.complete.obs")

#for dropdowns
occupations <- unique(employment_occupation$description)
industries <- unique(employment_region_industry$lmo_detailed_industry)
regions <- unique(employment_region_industry$bc_region)
```

Employment by Industry:Region
=====================================  

Inputs {.sidebar}
-------------------------------------

* If you hover over the plot on the left, a tooltip displays the industry and region.
* The further you are from the 45 degree line, the greater the difference between Stokes and Rich.


```{r}

selectInput(
  "industry",
  "choose an industry:",
  industries,
  selected = "Provincial and territorial public administration",
)

selectInput(
  "region",
  "choose a region:",
  regions,
  selected = "Lower Mainland-Southwest",
)

industry_fcasts <- reactive({
  employment_region_industry|>
    filter(lmo_detailed_industry==input$industry,
           bc_region==input$region)|>
    select(lmo_detailed_industry, bc_region, year, contains("employment"))|>
    pivot_longer(cols = contains("employment"))|>
    group_by(lmo_detailed_industry, bc_region, name)|>
    nest()|>
    mutate(cagr=map_dbl(data, get_cagr))|>
    unnest(data)
})

industry_lfs <- reactive({
  ind_lfs|>
    filter(lmo_detailed_industry==input$industry,
             bc_region==input$region)|>
      select(lmo_detailed_industry, bc_region, year, value)|>
    mutate(name="LFS",
           year=as.numeric(year))|>
    group_by(lmo_detailed_industry, bc_region)|>
    nest()|>
    mutate(cagr=map_dbl(data, get_cagr))|>
    unnest(data)
})
```

    
Column {data-width=500}
-------------------------------------
    
### Forecast comparison of all Industry:Region combinations
    
```{r}
rich_vs_stokes(employment_region_industry, richs_employment, stokes_employment, industry_employment_cor)
```
   
Column {data-width=500}
-------------------------------------
   
### Forecast comparison for the selected Industry:Region

```{r}
renderPlotly({
  plt <- ggplot()+
  geom_point(aes(year, 
                 employment, 
                 colour=series,
                 text=paste0(
                   "2021 Census Employment: ",
                   employment)), census_industry_region|>filter(lmo_detailed_industry==input$industry,
             bc_region==input$region))+ #need to fix this
  geom_line(aes(year, 
                value, 
                colour=name,
                group=name,
                text=paste0(str_to_title(str_replace(name,"_"," ")),
                   "\n Year: ",
                   year,
                   "\n Employment: ",
                   scales::comma(value, accuracy = 1),
                   "\n CAGR: ",
                   scales::percent(cagr, accuracy = .1))),
            bind_rows(industry_lfs(),industry_fcasts()))+
  labs(x=NULL, 
       y="Employment",
       colour=NULL,
       title=paste(input$industry, input$region, sep=": "))+
  scale_y_continuous(labels = scales::comma)
  
plotly::ggplotly(plt, tooltip="text")
})
```   
 

Employment by Occupation
=====================================

Inputs {.sidebar}
-------------------------------------

* If you hover over the plot on the left, a tooltip displays the occupation.
* The further you are from the 45 degree line, the greater the difference between Stokes and Rich.


```{r}
selectInput(
  "occupation",
  "choose an occupation:",
  occupations,
  selected = "Retail sales supervisors",
)

occupation_fcasts <- reactive({
  employment_occupation|>
    filter(description==input$occupation)|>
    select(noc_5, year, contains("employment"))|>
    pivot_longer(cols = contains("employment"))|>
    group_by(noc_5, name)|>
    nest()|>
    mutate(cagr=map_dbl(data, get_cagr))|>
    unnest(data)
})

occupation_lfs <- reactive({
  occ_lfs|>
    filter(description==input$occupation)|>
    mutate(name="LFS",
           year=as.numeric(year))|>
    group_by(noc_5, description)|>
    nest()|>
    mutate(cagr=map_dbl(data, get_cagr))|>
    unnest(data)
})
```


Column {data-width=500}
-------------------------------------
    
### Forecast comparison of all occupations
    
```{r}
rich_vs_stokes2(employment_occupation, richs_employment, stokes_employment, occupation_employment_cor)
```
   
Column {data-width=500}
-------------------------------------
   
### Forecast comparison for the selected occupation

```{r, fig.retina=2}
renderPlotly({
  plt <- ggplot()+
  geom_point(aes(year, 
                 employment, 
                 colour=series,
                 text=paste0(
                   "2021 Census Employment: ",
                   employment)), 
             census_occupation|>filter(description==input$occupation))+
  geom_line(aes(year, 
                value, 
                colour=name,
                group=name,
                 text=paste0(str_to_title(str_replace(name,"_"," ")),
                   "\n Year: ",
                   year,
                   "\n Employment: ",
                   scales::comma(value, accuracy = 1),
                   "\n CAGR: ",
                   scales::percent(cagr, accuracy = .1))), 
            bind_rows(occupation_lfs(),occupation_fcasts()))+
  labs(x=NULL, 
       y="Employment",
       colour=NULL,
       title=input$occupation)+
  scale_y_continuous(labels = scales::comma)

plotly::ggplotly(plt, tooltip="text")
})
```   
 
Demand
=====================================

Column {data-width=500}
-------------------------------------
    
### Expansion Demand by Industry:Region

```{r}
title <- jo_region_industry|>
  ungroup()|>
  summarize(rich=scales::comma(sum(richs_expansion, na.rm = TRUE)),
            stokes=scales::comma(sum(stokes_expansion, na.rm = TRUE)),
            correlation=round(cor(richs_expansion, stokes_expansion, method = "pearson", use="pairwise.complete.obs"),3))|>
  row_to_string()

demand_plot(jo_region_industry, richs_expansion, stokes_expansion, title)
```
    

### Replacement Demand by Industry:Region

```{r}
title <- jo_region_industry|>
  ungroup()|>
  summarize(rich=scales::comma(sum(richs_replacement, na.rm = TRUE)),
            stokes=scales::comma(sum(stokes_replacement, na.rm = TRUE)),
            correlation=round(cor(richs_replacement, stokes_replacement, method = "pearson", use="pairwise.complete.obs"),3))|>
  row_to_string()

demand_plot(jo_region_industry, richs_replacement, stokes_replacement, title)
```
 
Column {data-width=500}
-------------------------------------
   
### Expansion Demand by Occupation

```{r}
title <- jo_occupation|>
  ungroup()|>
  summarize(rich=scales::comma(sum(richs_expansion, na.rm = TRUE)),
            stokes=scales::comma(sum(stokes_expansion, na.rm = TRUE)),
            correlation=round(cor(richs_expansion, stokes_expansion, method = "pearson", use="pairwise.complete.obs"),3))|>
  row_to_string()

demand_plot2(jo_occupation, richs_expansion, stokes_expansion, title)
```

### Replacement Demand by Occupation

```{r}

title <- jo_occupation|>
  ungroup()|>
  summarize(rich=scales::comma(sum(richs_replacement, na.rm = TRUE)),
            stokes=scales::comma(sum(stokes_replacement, na.rm = TRUE)),
            correlation=round(cor(richs_replacement, stokes_replacement, method = "pearson", use="pairwise.complete.obs"),3))|>
  row_to_string()

demand_plot2(jo_occupation, richs_replacement, stokes_replacement, title)
```




