---
title: "Comparing Stokes to Rich"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
resource_files:
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(here)
library(plotly)
library(conflicted)
library(janitor)
library(readxl)
conflicts_prefer(dplyr::filter)
```

```{r}
#functions------------------------------
rich_vs_stokes <- function(tbbl, var1, var2, cor){
  plt <- ggplot(tbbl,  aes({{  var1  }},
                           {{  var2  }},
                           colour=year,
                           text=paste0(
                             "Region: ",
                             bc_region,
                             "\n Industry: ",
                             lmo_detailed_industry,
                             "\n Year: ",
                             year,
                             "\n Rich: ",
                             scales::comma({{  var1  }}, accuracy = 1),
                             "\n Stokes: ",
                             scales::comma({{  var2  }}, accuracy = 1)
                           )
  )
  )+
    geom_abline(slope = 1, intercept = 0, colour="white", lwd=2)+
    geom_point(size=.5)+
    scale_colour_viridis_c()+
    scale_x_continuous(trans="log10", labels=scales::comma)+
    scale_y_continuous(trans="log10", labels=scales::comma)+
    labs(title=paste("correlation: ",round(cor,3)))

  plotly::ggplotly(plt, tooltip="text")
}
rich_vs_stokes2 <- function(tbbl, var1, var2, cor){
  plt <- ggplot(tbbl,  aes({{  var1  }},
                           {{  var2  }},
                           colour=year,
                           text=paste0(
                             "NOC: ",
                             description,
                             "\n Year: ",
                             year,
                             "\n Rich: ",
                             scales::comma({{  var1  }}, accuracy = 1),
                             "\n Stokes: ",
                             scales::comma({{  var2  }}, accuracy = 1)
                           )
  )
  )+
    geom_abline(slope = 1, intercept = 0, colour="white", lwd=2)+
    geom_point(size=.5)+
    scale_colour_viridis_c()+
    scale_x_continuous(trans="log10", labels=scales::comma)+
    scale_y_continuous(trans="log10", labels=scales::comma)+
    labs(title=paste("correlation: ",round(cor,3)))

  plotly::ggplotly(plt, tooltip="text")
}
#read in the data----------------------

correct_industries <- read_csv(here("data","mapping", "tidy_2024_naics_to_lmo.csv"))|>
  select(naics, lmo_detailed_industry)|>
  select(lmo_detailed_industry)|>
  distinct()|>
  ungroup()

#my forecasts-----------------------------------
richs_forecast <- read_rds(here("out","richs_forecast.rds"))

richs_region_industry <- richs_forecast|>
  group_by(bc_region, lmo_ind_code, lmo_detailed_industry, year)|>
  summarise(richs_employment=sum(employment))

richs_occupation <- richs_forecast|>
  group_by(noc_5, year)|>
  summarise(richs_employment=sum(employment))

correct_regions <- richs_region_industry|>
  ungroup()|>
  select(bc_region)|>
  distinct()

#lfs data-----------------------
occ_lfs <-  read_excel(here("data", "Labour force status for 5 digit NOC 2014-2023.xlsx"), sheet = "Employed", skip=3)|>
    pivot_longer(cols = starts_with("2"), names_to = "year")|>
    clean_names()|>
  rename(description=class_title)|>
   mutate(noc_5=if_else(noc_5 %in% c("00011", "00012", "00013", "00014", "00015"), "00018", noc_5),
          description=if_else(noc_5=="00018","Senior managers - public and private sector", description))|>
  group_by(noc_5, description, year)|>
  summarize(value=sum(value, na.rm = TRUE))

ind_lfs_file <- here("data","Employment for 64 LMO Industries 2000-2023.xlsx")

ind_lfs <- excel_sheets(ind_lfs_file) %>%
  set_names() %>%
  map(~ read_excel(ind_lfs_file, sheet = .x, skip = 3))|>
  enframe()|>
  unnest(value)|>
  rename(bc_region=name)|>
  pivot_longer(cols=starts_with("2"), names_to = "year")|>
  clean_names()|>
  fuzzyjoin::stringdist_join(correct_industries)|>
  rename(lmo_detailed_industry=lmo_detailed_industry.y)|>
  fuzzyjoin::stringdist_join(correct_regions)|>
  rename(bc_region=bc_region.y)

#stokes data-------------------
stokes_employment_industry <- read_excel(here("data","stokes_data", "employment_industry.xlsx"), skip = 3)|>
  pivot_longer(cols=starts_with("2"), names_to = "year", values_to = "stokes_employment")|>
  clean_names()|>
  ungroup()|>
  rename(lmo_detailed_industry=industry)|>
  fuzzyjoin::stringdist_join(correct_industries)|>
  rename(lmo_detailed_industry=lmo_detailed_industry.y)|>
  select(bc_region=geographic_area, lmo_detailed_industry, year, stokes_employment)|>
  filter(bc_region!="British Columbia")|>
  mutate(year=as.numeric(year),
         bc_region=if_else(bc_region=="Mainland South West","Lower Mainland-Southwest", bc_region),
         bc_region=if_else(bc_region=="Vancouver Island Coast","Vancouver Island and Coast", bc_region)
  )|>
  fuzzyjoin::stringdist_full_join(correct_regions)|>
  rename(bc_region=bc_region.y)|>
  select(-bc_region.x)

stokes_employment_occupation <- read_excel(here("data","stokes_data", "employment_occupation.xlsx"), skip = 3)|>
  pivot_longer(cols=starts_with("2"), names_to = "year", values_to = "stokes_employment")|>
  clean_names()|>
  filter(geographic_area!="British Columbia")|>
  group_by(noc, description, year)|>
  summarize(stokes_employment=sum(stokes_employment))|>
  mutate(year=as.numeric(year),
         noc=str_sub(noc, 2))

#merging forecasts------------------------------

region_industry <- full_join(richs_region_industry, stokes_employment_industry)|>
  mutate(richs_employment=if_else(is.na(richs_employment), 0, richs_employment))

occupation <- full_join(richs_occupation, stokes_employment_occupation, by=c("noc_5"="noc", "year"="year"))|>
  mutate(richs_employment=if_else(is.na(richs_employment), 0, richs_employment))

#correlations between forecasts----------------------------

industry_employment_cor <- cor(region_industry$richs_employment, 
                               region_industry$stokes_employment, 
                               method = "pearson")

occupation_employment_cor <- cor(occupation$richs_employment, 
                               occupation$stokes_employment, 
                               method = "pearson",
                               use="pairwise.complete.obs")

#for dropdowns
occupations <- unique(occupation$description)
industries <- unique(region_industry$lmo_detailed_industry)
regions <- unique(region_industry$bc_region)
```

By Industry:Region
=====================================  

Inputs {.sidebar}
-------------------------------------

* If you hover over the plot on the left, a tooltip displays the industry and region.
* The further you are from the 45 degree line, the greater the difference between Stokes and Rich.
* Not much daylight between the forecasts.


```{r}

selectInput(
  "industry",
  "choose an industry:",
  industries,
  selected = "Provincial and territorial public administration",
)

selectInput(
  "region",
  "choose a region:",
  regions,
  selected = "Lower Mainland-Southwest",
)

industry_fcasts <- reactive({
  region_industry|>
    filter(lmo_detailed_industry==input$industry,
           bc_region==input$region)|>
    select(lmo_detailed_industry, bc_region, year, contains("employment"))|>
    pivot_longer(cols = contains("employment"))
})

industry_lfs <- reactive({
  ind_lfs|>
    filter(lmo_detailed_industry==input$industry,
             bc_region==input$region)|>
      select(lmo_detailed_industry, bc_region, year, value)|>
    mutate(name="LFS",
           year=as.numeric(year))
})
```

    
Column {data-width=500}
-------------------------------------
    
### Forecast comparison of all Industry:Region combinations
    
```{r}
rich_vs_stokes(region_industry, richs_employment, stokes_employment, industry_employment_cor)
```
   
Column {data-width=500}
-------------------------------------
   
### Forecast comparison for the selected Industry:Region

```{r}
renderPlotly({
  bind_rows(industry_lfs(),industry_fcasts())|>
  ggplot(aes(year, value, colour=name))+
  geom_line()+
  labs(x=NULL, 
       y="Employment",
       colour=NULL,
       title=paste(input$industry, input$region, sep=": "))+
  scale_y_continuous(labels = scales::comma)
})
```   
 

By Occupation
=====================================

Inputs {.sidebar}
-------------------------------------

* If you hover over the plot on the left, a tooltip displays the occupation.
* The further you are from the 45 degree line, the greater the difference between Stokes and Rich.
* The forecasts are substantially more different by occupation, and I would argue that my forecasts better reflect the LFS data.


```{r}
selectInput(
  "occupation",
  "choose an occupation:",
  occupations,
  selected = "Retail sales supervisors",
)

occupation_fcasts <- reactive({
  occupation|>
    filter(description==input$occupation)|>
    select(noc_5, year, contains("employment"))|>
    pivot_longer(cols = contains("employment"))
})

occupation_lfs <- reactive({
  occ_lfs|>
    filter(description==input$occupation)|>
    mutate(name="LFS",
           year=as.numeric(year))
})
```


Column {data-width=500}
-------------------------------------
    
### Forecast comparison of all occupations
    
```{r}
rich_vs_stokes2(occupation, richs_employment, stokes_employment, occupation_employment_cor)
```
   
Column {data-width=500}
-------------------------------------
   
### Forecast comparison for the selected occupation

```{r, fig.retina=2}
renderPlotly({
bind_rows(occupation_lfs(), occupation_fcasts())|>
  ggplot(aes(year, value, colour=name))+
  geom_line()+
  labs(x=NULL, 
       y="Employment",
       colour=NULL,
       title=input$occupation)+
  scale_y_continuous(labels = scales::comma)
})
```   
 



