---
title: "Margin manipulator"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(here)
library(vroom)
library(plotly)
library(shiny)
library(conflicted)
conflicts_prefer(vroom::cols)
conflicts_prefer(vroom::col_double)
conflicts_prefer(vroom::col_character)
conflicts_prefer(dplyr::filter)
#functions--------------------------------------
margin_plot <- function(tbbl, margin){
  tbbl|>
  ggplot()+
    geom_line(aes(year, employment, colour=series, lty=series))+
    facet_wrap(~ get(margin), scales = "free_y")+
    labs(x=NULL,
         y=NULL)+
    theme_minimal()+
    theme(text=element_text(size=13),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
}
get_total <- function(tbbl){
   sum(tbbl$employment)
 } 
get_cagr <- function(tbbl, horizon){
  max_year <- max(tbbl$year)
  start <- tbbl$employment[tbbl$year==max_year-horizon]
  end <-  tbbl$employment[tbbl$year==max_year]
  if_else(start==0, 1, (end/start)^(1/horizon)) #if start=0 the data is probably shitty: assume zero growth
}
occ_plot <- function(pg){
  occupation_cagrs <- post_mod_occupation()|>
    filter(page==pg)|>
    group_by(noc_5)|>
    nest()|>
    mutate(cagr=map_dbl(data, get_cagr, 10))|>
    select(noc_5, cagr)|>
    mutate(cagr=scales::percent(cagr-1, accuracy=.1))
  
  bind_rows(occupation_historic, post_mod_occupation(), pre_mod_occupation())|>
    filter(page==pg)|>
    full_join(occupation_cagrs)|>
    unite(noc_5, noc_5, cagr, sep=": ")|>
    ggplot()+
      geom_line(aes(year, employment, colour=series, lty=series))+
      facet_wrap(~noc_5, scales = "free_y")+
      labs(x=NULL,
           y=NULL)+
      theme_minimal()+
      theme(text=element_text(size=13),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks.y=element_blank())
}

# read in the data----------------------------
region_historic <- read_rds(here("out","historic_bc_region.rds"))
industry_historic <- read_rds(here("out", "historic_lmo_ind_code.rds"))
occupation_historic <- read_rds(here("out","historic_noc_5.rds"))|>
  arrange(noc_5)|>
  mutate(page=ntile(n=8))|>
  group_by(page)|>
  mutate(page_description=paste0(min(noc_5),"-",max(noc_5)))
noc_pages <- occupation_historic|> 
  select(noc_5, page)|>
  distinct()
mod_shares <-  read_rds(here("out","modified_shares","shares.rds"))
bc_fcast <- read_rds(here("out","bc_forecast.rds"))
regions <- unique(region_historic$bc_region)
industries <- unique(industry_historic$lmo_ind_code)
occupations <- unique(occupation_historic$noc_5)

```

Inputs {.sidebar}
=====================================


```{r}
# shiny inputs defined here
selectInput(
  "margin",
  "What margin do you want to manipulate?",
  c("bc_region", "noc_5", "lmo_ind_code"),
  "bc_region"
)
```

```{r}

dropdown <- reactive({
  if(input$margin=="bc_region"){
    regions
  }else if(input$margin=="noc_5"){
    occupations
  }else if(input$margin=="lmo_ind_code"){
    industries
  }else{
    stop("Error: select margin")
  }
})

renderUI({#need the renderUI because dropdown list is a reactive object
  selectInput('item', 
              'What item do you want to manipulate?', 
              dropdown()
              )
})

sliderInput("shift", "Shift up or down",
  min = .7, max = 1.3,
  value = 1, step = .01
)
sliderInput("slope", "Change slope",
  min = .95, max = 1.03,
  value = 1, step = .001
)
actionButton("save", "Save modified shares to disk")
```
 
```{r}
#reactive elements
shares <- reactive({
  mod_shares|>
      ungroup()|>
      mutate(multiplier=if_else(get(input$margin) == input$item, 
                                input$shift*input$slope^(year-min(year)), 
                                1))|>
    group_by(year)|>
    mutate(post_mod_share=pre_mod_share*multiplier,
           post_mod_share=post_mod_share/sum(post_mod_share))
})

pre_mod_forecast <- reactive({
 full_join(shares(), bc_fcast)|>
    mutate(employment=pre_mod_share*employment,
           series="Pre modification")
})

post_mod_forecast <- reactive({
  full_join(shares(), bc_fcast)|>
    mutate(employment=post_mod_share*employment,
           series="Post modification")
})

pre_mod_region <- reactive({
  pre_mod_forecast()|>
    group_by(year, bc_region, series)|>
    summarize(employment=sum(employment))
})

post_mod_region <- reactive({
  post_mod_forecast()|>
    group_by(year, bc_region, series)|>
    summarize(employment=sum(employment))
})

pre_mod_industry <- reactive({
  pre_mod_forecast()|>
    group_by(year, lmo_ind_code, series)|>
    summarize(employment=sum(employment))
})

post_mod_industry <- reactive({
  post_mod_forecast()|>
    group_by(year, lmo_ind_code, series)|>
    summarize(employment=sum(employment))
})

pre_mod_occupation <- reactive({
  pre_mod_forecast()|>
    group_by(year, noc_5, series)|>
    summarize(employment=sum(employment))|>
    full_join(noc_pages)
})

post_mod_occupation <- reactive({
  post_mod_forecast()|>
    group_by(year, noc_5, series)|>
    summarize(employment=sum(employment))|>
    full_join(noc_pages)
})

historic <- reactive({
  if(input$margin=="bc_region"){
    region_historic|>filter(bc_region==input$item)#|>mutate(series="historic")
  }else if(input$margin=="noc_5"){
    occupation_historic|>filter(noc_5==input$item)#|>mutate(series="historic")
  }else if(input$margin=="lmo_ind_code"){
    industry_historic|>filter(lmo_ind_code==input$item)#|>mutate(series="historic")
  }else{
    stop("error: historic")
  }
})

pre_mod <- reactive({
  if(input$margin=="bc_region"){
    pre_mod_region()|> filter(bc_region==input$item)
  }else if(input$margin=="noc_5"){
    pre_mod_occupation()|> filter(noc_5==input$item)
  }else if(input$margin=="lmo_ind_code"){
    pre_mod_industry()|> filter(lmo_ind_code==input$item)
  }else{
    stop("error: pre mod")
  }
})

post_mod <- reactive({
  if(input$margin=="bc_region"){
    post_mod_region()|> filter(bc_region==input$item)
  }else if(input$margin=="noc_5"){
    post_mod_occupation()|> filter(noc_5==input$item)
  }else if(input$margin=="lmo_ind_code"){
    post_mod_industry()|> filter(lmo_ind_code==input$item)
  }else{
    stop("error: post mod")
  }
})

observeEvent(input$save, {
  shares()|> #give the modified share tbbl same structure as pre_mod_share tbbl, and overwrite.
    select(-pre_mod_share, -multiplier)|>
    rename(pre_mod_share=post_mod_share)|>
    write_rds(here("out","modified_shares", "shares.rds"))
  tibble("margin"=input$margin, 
         "item"=input$item, 
         "shift"=input$shift, 
         "slope"=input$slope)|>
    write.table(here("out","change_log", "changes.csv"),
    append = TRUE,
    quote = TRUE,
    sep = ",",
    row.names = FALSE,
    col.names = !file.exists(here("out","change_log", "changes.csv"))
  )

})

```

Description
=====================================

### Intro

* For the BC LMO we are constrained by the Finance BC employment forecast (at least for the first 5 years)
* Given this constraint, in effect we are forecasting proportions of BC employment.
* There are 4 dimensions that we care about: Time, Region, Industry and Occupation.
* At a given point in time, one can visualize the remaining 3 dimensions as a cube, with sides representing the 7 regions, 64 industries, and 517 occupations, and cells representing the proportion of BC employment for that combination of region, industry and occupation.
* Note that this cube is extremely sparse: there are not many combinations of region, industry and occupation that have positive proportions.
* We are typically interested in looking at two dimensions at a time: e.g. industry employment over time, aggregating over the remaining two dimensions. 
* Our forecasting technique needs to overcome the limitations associated with the sparsity of the data.

### Approach taken:

*  Using LFS data from the years 2018,2019,2021,2022,2023 we calculate baseline proportions of BC employment based on every combination of region, industry and occupation.
*  We then aggregate the data by each of the above margins, and calculate the 5 year CAGR (2018--2023) for each item of each margin e.g. the recent historic growth rate for the Kootenays.
*  We then apply these marginal growth rates to adjust the baseline proportions.
*  For example, suppose that a region/industry/occupation combination accounted for 1% of the BC labour market over the years 2018,2019,2021,2022,2023.  Suppose that region's CAGR is 2%, the industry's CAGR is 3%, and the occupation's CAGR is -4%,
*  For the $t$th year of the forecast, the unadjusted proportion is $.01\times1.02^{t}\times1.03^{t}\times.96^{t}$.
*  Note that regarding the *change* in proportions over time, region, industry and occupation are assumed to be independent.
*  This is in contrast to the baseline levels, where the sparsity of the data indicates strong dependency between region, industry and occupation.
*  Note that these unadjusted proportions will not necessarily sum to 1: we adjust so they do so.

Manipulate
=====================================  

Row {data-height=900}
-------------------------------------

### 

```{r}
renderPlotly({
  req(input$margin)
  req(input$item)

  mod_cagr <- post_mod()|>
    get_cagr(10)
  mod_cagr <- scales::percent(mod_cagr-1, accuracy = .1) 

bind_rows(historic(), post_mod(), pre_mod())|>
  mutate(mod_cagr=mod_cagr)|>
  unite(facet_label, input$margin, mod_cagr, sep=": ")|>
  ggplot()+
    geom_line(aes(year, employment, colour=series, lty=series))+
    facet_wrap(~facet_label)+
    theme_minimal()
})
```

Row {data-height=100}
-------------------------------------
  

### Historic 5 year CAGR

```{r}
renderUI({
  req(input$margin)
  req(input$item)
  hist_cagr <- historic()|>
    get_cagr(5)
  scales::percent(hist_cagr-1, accuracy = .1)
})
```

### Base forecast 10 year CAGR
    
```{r}
renderUI({
  req(input$margin)
  req(input$item)
  base_cagr <- pre_mod()|>
    get_cagr(10)
  scales::percent(base_cagr-1, accuracy = .1)
}) 
```
 
### Modified 10 year CAGR
    
```{r}
renderUI({
  req(input$margin)
  req(input$item)
  mod_cagr <- post_mod()|>
    get_cagr(10)
  scales::percent(mod_cagr-1, accuracy = .1)
})
```
 
   
Region
====================================     

### Regional Employment

```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)

 regional_cagrs <- post_mod_region()|>
    group_by(bc_region)|>
    nest()|>
    mutate(cagr=map_dbl(data, get_cagr, 10))|>
    select(bc_region, cagr)|>
    mutate(cagr=scales::percent(cagr-1, accuracy=.1))
  
  bind_rows(region_historic, post_mod_region(), pre_mod_region())|>
    full_join(regional_cagrs)|>
    unite(bc_region, bc_region, cagr, sep=": ")|>
    margin_plot("bc_region")
})
```

Industry
=====================================  

### Industry Employment
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)

 industry_cagrs <- post_mod_industry()|>
    group_by(lmo_ind_code)|>
    nest()|>
    mutate(cagr=map_dbl(data, get_cagr, 10))|>
    select(lmo_ind_code, cagr)|>
    mutate(cagr=scales::percent(cagr-1, accuracy=.1))
  
  bind_rows(industry_historic, post_mod_industry(), pre_mod_industry())|>
    full_join(industry_cagrs)|>
    unite(lmo_ind_code, lmo_ind_code, cagr, sep=": ")|>
    margin_plot("lmo_ind_code")
})
``` 

   
`r unique(occupation_historic$page_description[occupation_historic$page==1])`
=====================================     

### 

```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(1)
})
``` 
    
`r unique(occupation_historic$page_description[occupation_historic$page==2])`
=====================================     

### 
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(2)
})
``` 
 
`r unique(occupation_historic$page_description[occupation_historic$page==3])`
=====================================     

### 
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(3)
})
```  

`r unique(occupation_historic$page_description[occupation_historic$page==4])`
=====================================     

### 
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(4)
})
```

`r unique(occupation_historic$page_description[occupation_historic$page==5])`
=====================================     

###  
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(5)
})
```

`r unique(occupation_historic$page_description[occupation_historic$page==6])`
=====================================     

###  
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(6)
})
```

`r unique(occupation_historic$page_description[occupation_historic$page==7])`
=====================================     

###  
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(7)
})
```

`r unique(occupation_historic$page_description[occupation_historic$page==8])`
=====================================     

###  
 
```{r, fig.retina=2}
renderPlotly({
  req(input$margin)
  req(input$item)
occ_plot(8)
})
```



